\pkg{casebase}: An Alternative Framework for Survival Analysis and Comparison of Event Rates

### 2022-11-22 analysis/paper/bhatnagar-turgeon-islam-hanley-saarela.Rmd
more updated captions

We then compare the risk estimation over the test set. The probabilities over time for each observation are averaged, resulting in the absolute risk curves shown in Figure \@ref(fig:cs3FinalBrier)A. We can see some minimal differences between the four models, with the Cox model giving the lowest estimates across follow-up-time. Note that the apparent smoothness of the Cox and penalized Cox curves is due to the large number of observations in the training set, which is used to derive the Breslow estimate of the baseline hazard. As described above, we compare the performance between the models by computing the Brier scores over time. In Figure \@ref(fig:cs3FinalBrier)B, we can see that the Brier score is larger for the Kaplan-Meier estimate than for the other three models. On the other hand, the differences between the three adjusted models are minimal.

@@

```{r cs3FinalBrier, echo = FALSE, eval = TRUE, fig.cap="Comparison of Cox regression (Cox), penalized Cox regression (Pen. Cox), penalized case-base sampling estimation (Pen. CB), and Kaplan-Meier (K-M). (A)  Probability of death as a function of follow-up time. (B) Brier score as a function of follow-up time, where a lower score corresponds to better performace. The unadjusted Kaplan-Meier gives the lowest risk estimates across follow-up-time and highest Brier scores, while we can see minimal differences between the three adjusted models across both metrics. Note that the apparent smoothness of the Cox and penalized Cox curves is due to the large number of observations in the training set, which is used to derive the Breslow estimate of the baseline hazard.", warning = FALSE}
# Combine scores
data_brier <- bind_rows(
  brierCoxKM$Brier$score %>% 
    mutate(model = as.character(model)) %>% 
    filter(model != "Null model"),
  brierPenalized$Brier$score %>% 
    mutate(model = as.character(model)) %>% 
    filter(model != "Null model")
  ) %>% 
  mutate(model = factor(model, levels = c("Cox", "Pen. Cox", "Pen. CB", "K-M")))

>>>>

We then compare the risk estimation over the test set. The predicted probabilities for each test set observation are averaged, resulting in the absolute risk curves shown in Figure \@ref(fig:cs3FinalBrier)A. The Kaplan-Meier curve is calculated on the test set only. We see minimal differences between the four approaches across follow-up-time. Note that the apparent smoothness of the Cox and penalized Cox curves is due to the large number of observations in the training set, which is used to derive the Breslow estimate of the baseline hazard. As described above, we compare the performance between the models by computing the Brier scores over time. In Figure \@ref(fig:cs3FinalBrier)B, we see that the Brier score is larger for the Kaplan-Meier estimate than for the other three models. On the other hand, the differences between the three adjusted models are minimal.

@@

```{r cs3FinalBrier, echo = FALSE, eval = TRUE, fig.cap="Comparison of Cox regression (Cox), penalized Cox regression (Pen. Cox), penalized case-base sampling estimation (Pen. CB), and Kaplan-Meier (K-M). (A) Probability of death as a function of follow-up time which is the average of the predicted probabilities for each test set observation. The Kaplan-Meier curve is calculated on the test set only. (B) Brier score as a function of follow-up time, where a lower score corresponds to better performance. We see minimal differences between the four approaches across follow-up-time for the absolute risk curves. Note that the apparent smoothness of the Cox and penalized Cox curves is due to the large number of observations in the training set, which is used to derive the Breslow estimate of the baseline hazard. We see that the Brier score is larger for the Kaplan-Meier estimate than for the other three models. On the other hand, the differences between the three adjusted models are minimal.", warning = FALSE}
# Combine scores
data_brier <- bind_rows(
  brierCoxKM$Brier$score %>% 
    mutate(model = as.character(model)) %>% 
    filter(model != "Null model"),
  brierPenalized$Brier$score %>% 
    mutate(model = as.character(model)) %>% 
    filter(model != "Null model")
  ) %>% 
  mutate(model = factor(model, levels = c("Cox", "Pen. Cox", "Pen. CB", "K-M")))

### 2022-11-22 analysis/paper/bhatnagar-turgeon-islam-hanley-saarela.tex
more updated captions

We then compare the risk estimation over the test set. The probabilities over time for each observation are averaged, resulting in the absolute risk curves shown in Figure \ref{fig:cs3FinalBrier}A. We can see some minimal differences between the four models, with the Cox model giving the lowest estimates across follow-up-time. Note that the apparent smoothness of the Cox and penalized Cox curves is due to the large number of observations in the training set, which is used to derive the Breslow estimate of the baseline hazard. As described above, we compare the performance between the models by computing the Brier scores over time. In Figure \ref{fig:cs3FinalBrier}B, we can see that the Brier score is larger for the Kaplan-Meier estimate than for the other three models. On the other hand, the differences between the three adjusted models are minimal.

\begin{figure}[ht]
\includegraphics[width=\textwidth,keepaspectratio=true]{./cs3FinalBrier-1} \caption{Comparison of Cox regression (Cox), penalized Cox regression (Pen. Cox), penalized case-base sampling estimation (Pen. CB), and Kaplan-Meier (K-M). (A)  Probability of death as a function of follow-up time. (B) Brier score as a function of follow-up time, where a lower score corresponds to better performace. The unadjusted Kaplan-Meier gives the lowest risk estimates across follow-up-time and highest Brier scores, while we can see minimal differences between the three adjusted models across both metrics. Note that the apparent smoothness of the Cox and penalized Cox curves is due to the large number of observations in the training set, which is used to derive the Breslow estimate of the baseline hazard.}\label{fig:cs3FinalBrier}
\end{figure}

>>>>

We then compare the risk estimation over the test set. The predicted probabilities for each test set observation are averaged, resulting in the absolute risk curves shown in Figure \ref{fig:cs3FinalBrier}A. The Kaplan-Meier curve is calculated on the test set only. We see minimal differences between the four approaches across follow-up-time. Note that the apparent smoothness of the Cox and penalized Cox curves is due to the large number of observations in the training set, which is used to derive the Breslow estimate of the baseline hazard. As described above, we compare the performance between the models by computing the Brier scores over time. In Figure \ref{fig:cs3FinalBrier}B, we see that the Brier score is larger for the Kaplan-Meier estimate than for the other three models. On the other hand, the differences between the three adjusted models are minimal.

\begin{figure}[ht]
\includegraphics[width=\textwidth,keepaspectratio=true]{./cs3FinalBrier-1} \caption{Comparison of Cox regression (Cox), penalized Cox regression (Pen. Cox), penalized case-base sampling estimation (Pen. CB), and Kaplan-Meier (K-M). (A) Probability of death as a function of follow-up time which is the average of the predicted probabilities for each test set observation. The Kaplan-Meier curve is calculated on the test set only. (B) Brier score as a function of follow-up time, where a lower score corresponds to better performance. We see minimal differences between the four approaches across follow-up-time for the absolute risk curves. Note that the apparent smoothness of the Cox and penalized Cox curves is due to the large number of observations in the training set, which is used to derive the Breslow estimate of the baseline hazard. We see that the Brier score is larger for the Kaplan-Meier estimate than for the other three models. On the other hand, the differences between the three adjusted models are minimal.}\label{fig:cs3FinalBrier}
\end{figure}
